---
title: 'Reproducible Research: Peer Assessment 1'
author: "João Pedro"
date: "25/04/2020"
output: html_document
keep_md: TRUE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Loading the data
```{r, results='hide'}
library(readr)
setwd("C:/Users/mcjpe/Desktop/João Pedro/Ciência de Dados/Tarefa 1 - Pesquisa Reprodusível")
data <- read_csv("activity.csv", na = 'NA')
```

## 2. Average of total steps taken per day
Firt, remove the NA's values from the data:
```{r, results='hide'}
data0 <- data[(!is.na(data$steps)),]
```
With this new data, calculate the sum of all steps per day and make a histogram of this results:
```{r}
stpd <- with(data0, tapply(steps, date, sum))
summary(stpd)
#png("Plot1.png")
hist(stpd, 
     main = 'Steps Taken Per Day', 
     xlab = 'Steps', 
     col = 'red', 
     breaks = length(stpd)/2)
#dev.off()
```

## 3. Daily activity average pattern
Initially, calculate the average steps per time interval using the data without NA's and create a new DataFrame with the results:
```{r}
step <- (with(data0, tapply(steps, interval, mean)))
name <- names(step)
stpm <- as.data.frame(step)
stpm$interval <- name
```
After that, calculate the maximum average value and get the time interval that's occur:
```{r}
maxm <- stpm[max(stpm$step),]
maxm <- as.numeric(maxm$interval)
maxm
```
Finally, plot the results:
```{r}
#png("Plot2.png")
plot(stpm$interval, stpm$step,
     main = 'Average Steps Taken Per 5 Minutes Interval', 
     xlab = 'Interval', 
     ylab = 'Steps', 
     type = 'l',
     lwd = 2,
     col = 'blue')
#dev.off()
```

## 4. Imputing missing values
Find the NA's values and insert the average steps in the correspond interval, after, create a new DataFrame:
```{r}
narow <- data[is.na(data$steps),]
narow <- split(narow, narow$date)
for (a in 1:length(narow)) {narow[[a]]$steps <- step}
narow <- do.call(rbind,narow)
data1 <- data
data1[is.na(data1$steps),] <- narow
```
With the new DataFrame, calculate the sum of the steps per day and plot the results. For compare the data with NA's values and without, make a panel plot:
```{r}
stpd1 <- with(data1, tapply(steps, date, sum))
summary(stpd1)
#png("Plot3.png")
par(mfcol = c(1,2))
hist(stpd, 
     main = 'Steps Taken Per Day \n(without NA)', 
     xlab = 'Steps', 
     col = 'red', 
     breaks = length(stpd)/2)
hist(stpd1, 
     main = 'Steps Taken Per Day \n(with NA)', 
     xlab = 'Steps', 
     col = 'green', 
     breaks = length(stpd1)/2)
#dev.off()
```

## 5. Differences in activity patterns between weekdays and weekends
First, separe the weekdays and the weekends in two DataFrames:
```{r}
data2 <- data1
data2$week <- as.character(weekdays(data2$date, abbreviate = TRUE))
data2[(data2$week == c("dom", "sab")),]$week <- "Weekend"
data2[(data2$week != "Weekend"),]$week <- "Weekday"
data2 <- split(data2, data2$week)
```
After that, do the same analysis did in step 3 for the two new DataFrames and plot the results in a panel plot:
```{r}
stwk <- with(data2[[1]], tapply(steps, interval, mean))
nmwk <- names(stwk)
stwk <- as.data.frame(stwk)
stwk$interval <- nmwk

stwe <- with(data2[[2]], tapply(steps, interval, mean))
nmwe <- names(stwe)
stwe <- as.data.frame(stwe)
stwe$interval <- nmwe

#png("Plot4.png")
par(mfrow = c(2,1))
plot(stwk$interval, stwk$stwk,
     main = 'Average Steps Taken Per 5 Minutes Interval (Weekday)', 
     xlab = 'Minutes', 
     ylab = 'Average Steps', 
     type = 'l',
     lwd = 2,
     col = 'cyan')
plot(stwe$interval, stwe$stwe,
     main = 'Average Steps Taken Per 5 Minutes Interval (Weekend)', 
     xlab = 'Minutes', 
     ylab = 'Average Steps', 
     type = 'l',
     lwd = 2,
     col = 'orange')
#dev.off()
```
